---
title: "Homework 3"
author: "Caleb Neely"
format: pdf
---

### Set up
```{r setup, include=FALSE}
# load package
library(tidyverse)
library(rstatix)

# read in dataset
temp_kelp <- read.csv("temp-kelp.csv")
```

## Problem 1

### a) An Appropriate Test

To determine the strength of the relationship between ocean temperature (°C) and giant kelp frond elongation rate (cm day<sup>-1</sup>), appropriate tests would be a **Pearson correlation** and a **Spearman rank correlation**. Pearson correlation measures the strength and direction of a **linear relationship** between the two continuous variables, while Spearman rank correlation measures the strength and direction of a **monotonic relationship** based on ranked data and is less sensitive to non-normal data and outliers.

### b) Visualization

```{r}
ggplot(temp_kelp, 
       aes(x = temp_c, 
           y = kelp_elong)
       ) +

  geom_point(
    color = "darkgreen",
    size = 1
  ) +

  geom_smooth(
    method = "lm",
    se = TRUE,
    color = "steelblue"
  ) +

  labs(
    x = "Ocean temperature (°C)",
    y = expression("Giant kelp elongation rate (cm day"^-1*")"),
    title = "Relationship between ocean temperature and giant kelp growth"
  ) +

  theme_bw()
```


### c) Check Assumptions And Run Test



















